Omics-based Single-step Trait Predition

### 2017-03-14 reports/index.Rmd
Methods: Init (single step) prediction models

---
title: Single-step Hybrid Predition
author:
  - name: Matthias Westhues
    email: matthias.westhues@uni-hohenheim.de
    affiliation: University of Hohenheim
  - name: Claas Heuer
    email: cheuer@tierzucht.uni-kiel.de
    affiliation: University of Kiel
  - name: Georg Thaller
    email: thaller@tierzucht.uni-kiel.de
    affiliation: University of Kiel
  - name: Rohan Fernando
    email: rohan@iastate.edu
    affiliation: Iowa State University
  - name: Albrecht E. Melchinger
    email: melchinger@uni-hohenheim.de
    affiliation: University of Hohenheim
    footnote: Corresponding Author
address:
  - code: University of Hohenheim
    address: Institute of Plant Breeding, Seed Science and Population Genetics, Fruwirthstr. 21, 70599 Stuttgart, Germany
  - code: University of Kiel
    address: Institut of Animal Breeding and Husbandry, Hermann-Rodewald-Straße 6, 24118 Kiel, Germany
  - code: Iowa State University
    address: Department of Animal Science, 239 Kildee Hall, 50011 Ames, Iowa, U.S.A.
date: "`r Sys.Date()`"
knit: "bookdown::render_book"
output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
bibliography: ["library.bib", "cheuer.bib"]
abstract: |
  This is the abstract.

>>>>

---
title: Omics-based Single-step Trait Predition
author:
  - name: Matthias Westhues
    email: matthias.westhues@uni-hohenheim.de
    affiliation: University of Hohenheim
  - name: Claas Heuer
    email: cheuer@tierzucht.uni-kiel.de
    affiliation: University of Kiel
  - name: Georg Thaller
    email: thaller@tierzucht.uni-kiel.de
    affiliation: University of Kiel
  - name: Rohan Fernando
    email: rohan@iastate.edu
    affiliation: Iowa State University
  - name: Albrecht E. Melchinger
    email: melchinger@uni-hohenheim.de
    affiliation: University of Hohenheim
    footnote: Corresponding Author
address:
  - code: University of Hohenheim
    address: Institute of Plant Breeding, Seed Science and Population Genetics, Fruwirthstr. 21, 70599 Stuttgart, Germany
  - code: University of Kiel
    address: Institut of Animal Breeding and Husbandry, Hermann-Rodewald-Straße 6, 24118 Kiel, Germany
  - code: Iowa State University
    address: Department of Animal Science, 239 Kildee Hall, 50011 Ames, Iowa, U.S.A.
date: "`r Sys.Date()`"
knit: "bookdown::render_book"
output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
bibliography: ["~/Dropbox/Paper/library.bib"]
abstract: |
  This is the abstract.

### 2017-03-14 reports/intro.Rmd
Methods: Init (single step) prediction models

# Introduction
The integration of different sources of information for the prediction of 
breeding or production values is one of the major challenges in current genomic 
prediction research.
In the early years of genomic prediction, a 2-step procedure was the most 
intuitive choice for establishing a new selection scheme.
The first consists of conventionally estimating breeding values based on 
pedigree relationships and therefore the phenotypes of relatives.
Individuals with accurate breeding values out of the first step (i.e. high 
number of offspring), enter the second step in which SNP genotypes are regressed 
against the conventional breeding values.
Those markers can then be used to predict genomic breeding values for solely 
genotyped and potentially very young individuals.
Considering the two sources of information &mdash; pedigree and high density SNP 
genotypes &mdash; the process can be described as follows:

@@

Consider now the situation in which only a subset of the population has 
transcriptomic information but all have SNP genotypes.
We can take a similar route as in @fernando_class_2014 and impute missing
mRNA data.
The difference here is that &mdash; instead of predicting SNP covariates using 
pedigree relationship &mdash; we impute mRNA covariates using SNP or genomic 
relationships.
Let the subscript $1$ denote individuals for which SNP genotypes but no mRNA 
data are available.
Individuals with subscript $2$ have both, SNP and mRNA information.
The covariates in $\mathbf{W}$ are centered.
The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
expectation given $\mathbf{g}_2$ and a residual error term:
\begin{align}
	(\#eq:mrna1)
	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
	&= \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
\end{align}

The covariance matrix of $\boldsymbol{\epsilon}$ is $\mathbf{G}_{11} - \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{G}_{21} = (\mathbf{G}^{11})^{-1}$
[@legarra_relationship_2009].

@@

Following the notation in @fernando_class_2014 the phenotypes can be untangled like this:

>>>>

# Introduction
The integration of different sources of information for the prediction of 
breeding or production values is one of the major challenges in current genomic 
prediction research.
In the early years of genomic prediction, a 2-step procedure was the most 
intuitive choice for establishing a new selection scheme.
The first consists of conventionally estimating breeding values based on 
pedigree relationships and therefore the phenotypes of relatives.
Individuals with accurate breeding values out of the first step (i.e. high 
number of offspring), enter the second step in which conventional breeding 
values are regressed onto SNP genotypes.
Those markers can then be used to predict genomic breeding values for solely 
genotyped and potentially very young individuals.
Considering the two sources of information &mdash; pedigree and high density SNP 
genotypes &mdash; the process can be described as follows:

@@

Consider now the situation in which only a subset of the population has 
transcriptomic information but all have SNP genotypes.
We can take a similar route as in @Fernando2014 and impute missing
mRNA data.
The difference here is that &mdash; instead of predicting SNP covariates using 
pedigree relationship &mdash; we impute mRNA covariates using SNP or genomic 
relationships.
Let the subscript $1$ denote individuals for which SNP genotypes but no mRNA 
data are available.
Individuals with subscript $2$ have both, SNP and mRNA information.
The covariates in $\mathbf{W}$ are centered.
The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
expectation given $\mathbf{g}_2$ and a residual error term:
\begin{align}
	(\#eq:mrna1)
	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
	&= \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
\end{align}

The covariance matrix of $\boldsymbol{\epsilon}$ is $\mathbf{G}_{11} - \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{G}_{21} = (\mathbf{G}^{11})^{-1}$
[@Legarra2009].

@@

Following the notation in @Fernando2014 the phenotypes can be untangled like this:

### 2017-03-14 reports/methods.Rmd
Methods: Init (single step) prediction models

### Hybrid data
The maize hybrid data set comprises 1,521 hybrids produced in 16 factorial
mating designs between 142 Dent and 103 Flint parent lines [@Westhues2017]. 
Best linear unbiased estimates (BLUEs) were computed across all factorials and 
three or more agro-ecologically diverse locations across Germany for six 
agronomically important traits.
All 245 parent lines have pedigree records back to the generation of their 
grandparents or deeper [@Westhues2017] and were genotyped using the Illumina 
SNP BeadChip MaizeSNP50 [@Ganal2011].
After filtering for minor allele frequency ($ \geq 5\%$), heterozygosity rate
($\geq 5\%$) and call frequency ($\geq 95\%$), missing genotypes were imputed 
using the Beagle software [@Browning2009].
The final number of polymorphic marker loci was 7,013 for parental Dent and
6,212 for parental Flint lines, respectively.
Gene expression data for seedlings seven days after sowing were obtained for a 
subset of 60 parental Dent and 43 parental Flint lines through two-color
hybridizations using a custom 2K microarray (GPL22267) [@Westhues2017].
After applying established normalization procedures [@Smyth2003,Ritchie2007], 
gene expression BLUEs were computed for 1,323 transcripts in reference to 
established protocols [@Symth2003,Ritchie2007,Frisch2010] using the R-package 
*limma* [@Ritchie2015].

#### Diversity panel
The maize inbred line data set originally comprised 513 maize lines representing 
the global maize diversity and was reduced to the set of tropical and
subtropical lines ($n = 211$) given that it is the largest of the four
pre-classified subgroups [@Yang2014].
All inbred lines were evaluated in five Chinese environments ranging from
$18^{\circ}$ to $30^{\circ}$N and from $102^{\circ}$ to $110^{\circ}$E
[@Yang20114].
Best linear unbiased predictors (BLUPs) were calculated for 17 traits of which 
we will evaluate six in our prediction models.
MaizeSNP50 BeadChip-data were available for the entire set of lines from the
diversity panel.
Additionally, RNA sequencing was performed for 368 lines on young seedlings 15 
days after sowing.
By exploting identify by descent (IBD) between 49,728 SNPs on the BeadChip 
overlapping with the RNA-seq data, 556,809 high quality SNPs could be inferred 
for these lines [@Fu2013,Li2013].
For the remaining 145 maize lines without RNA-seq data, high density markers
were inferred via projection of IBD regions onto the BeadChip, using this core 
set of markers [@Yang2014].
The same SNP quality checks as for the parental inbred lines of the hybrids
were applied to the 211 inbred lines from the tropical/subtropical subset,
yielding 37,760 SNPs.
Gene expression data were normalized using a normal quantile transformation to
satisfy modelling assumptions [@Fu2013].
A subset of 28,850 annotated genes was available for 149 out of the 211
genotypes and was kept for further analyses [@Li2013].

@@

### Kernels
BLUP models are computationally attractive when the number of features $p$
exceeds the number of genotypes $n$ and are equivalent to a selection index
when fixed effect have been accomodated by the dependent variable
(@Mrode2014, pp. 34, 311, 312).
Depending on the data set, up to three predictors were available for agronomic
trait predictions, namely pedigree data (P), genomic data (G) and
transcriptompic data (T).
The corresponding feature matrix of the $l$-th group of inbred lines and the 
$o$-th predictor ($\mathbf{W_{lo}}$) has dimensions $n \times p$ where $n$
pertains to the number of gneotypes in the $l$-th group and $p$ pertains to 
the number of features.
For the diversity panel maize lines, $n$ corresponds to the number of genotypes
whereas, in the case of the hybrid data, $n$ corresponds to the number of
parent lines in the corresponding heterotic group.
In generating kernels for each predictor, all features in $\mathbf{W_{lo}}$ were
centered and standardized to unit variance.
Then, each kernel can be defined as

$$
\begin{equation} \label{eq:GenomicRelationship}
  \mathbf{G}_{lo} = \frac{1}{W_{lo}} \mathbf{W}_{lo} \mathbf{W}_{lo}^{\top},
\end{equation}
$$
where $W$ denotes the number of features [@VanRaden2008].
In the case of pedigree data (P), coancestry coefficients were used directly
for $\mathbf{G}_{lo}$.

### Models

### Imputation

### Leave-one-out cross-validation
In the case of the diversity panel maize inbred lines, leave-one-out cross
validation (LOOCV) was done by simply using a single genotype as a hold-out
sample, which will subsequently be predicted, while using all other inbred 
lines for model training.
This process is repeated until all inbred lines have been used once for testing
and $n - 1$ times for model training.

For the hybrid data, LOOCV was carried out as follows:
Let $D$ and $F$ denote the set of parental inbred lines from the Dent and the 
Flint group, respectively.
Further, let $H \cap [D \times F]$ denote the set of hybrids from crosses 
between $D$ and $F$.
The training set ($H_{TRN}$) for the hybrid to be predicted ($H_{ij}$) --- 
where $i \in D_{TRN}$ and $j \in F_{TRN}$ --- was assembled as 
$H_{TRN} = [H \cap (D_{TRN}^{C} \times F_{TRN}^{C})]$, where 
$D_{TRN}^{C} = D \setminus D_{TRN}$ and $F_{TRN}^{C} = F \setminus F_{TRN}$.

>>>>

### Hybrid data
The maize hybrid data set comprises 1,521 hybrids produced in 16 factorial
mating designs between 142 Dent and 103 Flint parent lines [@Westhues2017]. 
Best linear unbiased estimates (BLUEs) were computed across all factorials and 
three or more agro-ecologically diverse locations across Germany for six 
agronomically important traits.
All 245 parent lines have pedigree records back to the generation of their 
grandparents or deeper [@Westhues2017] and were genotyped using the Illumina 
SNP BeadChip MaizeSNP50 [@Ganal2011].
After filtering for minor allele frequency ($\geq 5$%), heterozygosity rate
($\geq 5$%) and call frequency ($\geq 95$%), missing genotypes were imputed 
using the Beagle software [@Browning2009].
The final number of polymorphic marker loci was 7,013 for parental Dent and
6,212 for parental Flint lines, respectively.
Gene expression data for seedlings seven days after sowing were obtained for a 
subset of 60 parental Dent and 43 parental Flint lines through two-color
hybridizations using a custom 2K microarray (GPL22267) [@Westhues2017].
After applying established normalization procedures [@Smyth2003;@Ritchie2007], 
gene expression BLUEs were computed for 1,323 transcripts in reference to 
established protocols [@Smyth2003;@Ritchie2007;@Frisch2010] using the R-package 
*limma* [@Ritchie2015a].

#### Diversity panel
The maize inbred line data set originally comprised 513 maize lines representing 
the global maize diversity and was reduced to the set of tropical and
subtropical lines ($n = 211$) given that it is the largest of the four
pre-classified subgroups [@Yang2014].
All inbred lines were evaluated in five Chinese environments ranging from
$18^{\circ}$ to $30^{\circ}$N and from $102^{\circ}$ to $110^{\circ}$E
[@Yang2014].
Best linear unbiased predictors (BLUPs) were calculated for 17 traits of which 
we will evaluate six in our prediction models.
MaizeSNP50 BeadChip-data were available for the entire set of lines from the
diversity panel.
Additionally, RNA sequencing was performed for 368 lines on young seedlings 15 
days after sowing.
By exploting identify by descent (IBD) between 49,728 SNPs on the BeadChip 
overlapping with the RNA-seq data, 556,809 high quality SNPs could be inferred 
for these lines [@Fu2013;@Li2013].
For the remaining 145 maize lines without RNA-seq data, high density markers
were inferred via projection of IBD regions onto the BeadChip, using this core 
set of markers [@Yang2014].
The same SNP quality checks as for the parental inbred lines of the hybrids
were applied to the 211 inbred lines from the tropical/subtropical subset,
yielding 37,760 SNPs.
Gene expression data were normalized using a normal quantile transformation to
satisfy modelling assumptions [@Fu2013].
A subset of 28,850 annotated genes was available for 149 out of the 211
genotypes and was kept for further analyses [@Li2013].

@@

### Kernels
BLUP models are computationally attractive when the number of features $p$
exceeds the number of genotypes $n$ and are equivalent to a selection index
when fixed effect have been accomodated by the dependent variable
(@Mrode2014, pp. 34, 311, 312).
Depending on the data set, up to three predictors were available for agronomic
trait predictions, namely pedigree data (P), genomic data (G) and
transcriptompic data (T).
The corresponding feature matrix of the $l$-th group of inbred lines and the 
$o$-th predictor ($\mathbf{W_{lo}}$) has dimensions $n \times p$ where $n$
pertains to the number of genotypes in the $l$-th group and $p$ pertains to 
the number of features.
For the diversity panel maize lines, $n$ corresponds to the number of genotypes
whereas, in the case of the hybrid data, $n$ corresponds to the number of
parent lines in the corresponding heterotic group.
In generating kernels for each predictor, all features in $\mathbf{W_{lo}}$ were
centered and standardized to unit variance.
Then, each kernel can be defined as

\begin{equation} \label{eq:GenomicRelationship}
  \mathbf{G}_{lo} = \frac{1}{W_{lo}} \mathbf{W}_{lo} \mathbf{W}_{lo}^{\top},
\end{equation}

where $W$ denotes the number of features [@VanRaden2008].
In the case of pedigree data (P), coancestry coefficients were used directly
for $\mathbf{G}_{lo}$.

### Models
The universal model for breeding values was:

\begin{equation} \label{eq:GBLUPModel}
  \mathbf{y} = \mu + 
  \sum_{l=1}^{L} \mathbf{Z}_{lo} \mathbf{g}_{lo} +
  \mathbf{\epsilon},
\end{equation}

where $\mathbf{y}$ is the vector of observed inbred line or hybrid performance,
respectively, $\mu$ is the fixed model intercept, $\mathbf{Z}_{lo}$ is the 
corresponding design matrix associating the random genotype effects (*i.e.*, 
breeding values) of the lines in the $l$-th inbred line group and the $o$-th 
predictor type with $\mathbf{y}$.
Note that, in the case of hybrid data, $\mathbf{Z}_{lo}$ associates the random
general combining ability effects in the parental inbred lines (either Dent or
Flint, as expressed by $l$) with the vector of observed hybrid performance
$\mathbf{y}$.
The random effects ($\mathbf{g}_{lo}$) have expectation zero and variance equal 
to $\mathbf{G}_{lo} \sigma^{2}_{{u}^{lo}}$ and $\mathbf{I} \sigma^2_{\epsilon}$ 
for the residual error.
All prediction models were implemented using the R-package *BGLR* [@Perez2014].

### Single Step Prediction

The breeding values in a prediction model are in general given by:
\begin{equation} \label{eq:mrnaebv}
	\mathbf{\tilde{g}} = \mathbf{W}\boldsymbol{\hat{\alpha}},
\end{equation}

where $\boldsymbol{\hat{\alpha}}$ is the solution to a ridge regression model,
which is equivalent to the BLUP model.

Consider now the situation in which one predictor is complete in the sense that
it contains information on the whole population but a second predictor is
incomplete and covers only a subset of the population.
We can take a similar route as in @Fernando2014 and impute covariates of the 
incomplete predictor by using covariates of the complete predictor.
Let the subscript $1$ denote individuals covered only by the complete
predictor whereas individuals covered by both, the complete and the
incomplete predictor, are indicated by the subscript $2$.
Furhter, the covariates in $\mathbf{W}$ are centered.
The vector $\mathbf{g}_1$ can be written as the sum of the conditional 
expectation given $\mathbf{g}_2$ and a residual error term:
\begin{align} \label{eq:mrna1}
	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) + \boldsymbol{\epsilon} \\
	&= \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}} + (\mathbf{g_1} - \mathbf{G_{12}}\mathbf{G_{22}}^{-1}\mathbf{W_2}\boldsymbol{\hat{\alpha}}) \\
	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon}
\end{align}

The covariance matrix of $\boldsymbol{\epsilon}$ is $\mathbf{G}_{11} - \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{G}_{21} = (\mathbf{G}^{11})^{-1}$
[@Legarra2009].

The structure of the residual imputation/prediction error is known and can 
therefore be modelled.
The covariates not covered by the incomplete predictor can be predicted using 
the expectation of a multivariate normal random vector given correlated 
observations, which are related to the Best Linear Predictor.

The general model for the single step procedure can then be written as

\begin{equation}
  (\#eq:single-step-model)
\mathbf{y} = \mathbf{Xb} + \mathbf{W} \boldsymbol{\alpha} + \mathbf{U}_{Full} \boldsymbol{\epsilon}_{Full} + \mathbf{e},
\end{equation}

with

\begin{equation}
 (\#eq:single-step-submatrices)
\mathbf{X} = 
\begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix},
 \mathbf{W} = 
\begin{bmatrix}
  Z_1\hat{\mathbf{W}_1} \\
  \mathbf{W}_2 
 \end{bmatrix},
 \mathbf{U}_{Full} = 
\begin{bmatrix}
  Z_1 \\
  0 
 \end{bmatrix}
\end{equation}

Following the notation in @Fernando2014 the phenotypes can be untangled like this:

\begin{align}
 (\#eq:entangled-augmented-single-step-model)
\begin{bmatrix}
  y_1 \\
  y_2 
 \end{bmatrix}
& =
 \begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_1 & 0 \\
  0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  g_1 \\
  g_2 
 \end{bmatrix}
  + \mathbf{e} \\
    & = 
 \begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_1 & 0 \\
  0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  \mathbf{G}_{12}\mathbf{G}_{22}^{-1}\mathbf{W}_2\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{Full}  \\
  \mathbf{W}_2\boldsymbol{\alpha} \\
 \end{bmatrix}
  + \mathbf{e} \\
    & = 
 \begin{bmatrix}
  X_1 \\
  X_2 
 \end{bmatrix}
 \boldsymbol{\beta} + 
 \begin{bmatrix}
  Z_1 & 0 \\
  0 & Z_2 
 \end{bmatrix}
\begin{bmatrix}
  \hat{\mathbf{W}_1}\boldsymbol{\alpha} + \boldsymbol{\epsilon}_{Full} \\
  \mathbf{W}_2\boldsymbol{\alpha} \\
 \end{bmatrix}
  + \mathbf{e} \\
\end{align}

The breeding values from that model are:

\begin{equation}
 (\#eq:breeding-values)
\tilde{\mathbf{g}} = 
 \begin{bmatrix}
  \hat{\mathbf{W}_1} \\
  \mathbf{W}_2 
 \end{bmatrix}
 \hat{\boldsymbol{\alpha}}
 + 
 \begin{bmatrix}
  \mathbf{Z}_1 \\
  0
 \end{bmatrix}
 \hat{\boldsymbol{\epsilon}}_{Full} 
\end{equation}

Our final model looks like this:

\begin{align}
 (\#eq:final-model)
\mathbf{y} &= \mathbf{Xb} +
\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{W}_{l} \boldsymbol{\alpha}_{l} + 
\sum_{l=1}^{L} \mathbf{Z}_{l}\mathbf{U}_{l{Full}} \boldsymbol{\epsilon}_{l{Full}} +
\mathbf{e}.
\end{align}

For the hybrid data, the obtained GCA effects for the inbred lines represent 
half their breeding values.
The predicted agronomic performance is:

\begin{equation}
 (\#eq:predicted-performance)
\tilde{\mathbf{h}} = \sum_{l=1}^{L} \mathbf{Z}_{l}\tilde{\mathbf{g}}_{l} 
\end{equation}

### Model validation
For the validation of our predictions, we employed leave-one-out
cross-validation (LOOCV) routines.
In the case of the diversity panel maize inbred lines, LOOCV was done by simply 
using a single genotype as a hold-out sample, which will subsequently be 
predicted, while using all other inbred lines for model training.
This process is repeated until all inbred lines have been used once for testing
and $n - 1$ times for model training.

For the hybrid data, LOOCV was carried out as follows:
Let $D$ and $F$ denote the set of parental inbred lines from the Dent and the 
Flint group, respectively.
Further, let $H \cap [D \times F]$ denote the set of hybrids from crosses 
between $D$ and $F$.
The training set ($H_{TRN}$) for the hybrid to be predicted ($H_{ij}$) --- 
where $i \in D_{TRN}$ and $j \in F_{TRN}$ --- was assembled as 
$H_{TRN} = [H \cap (D_{TRN}^{C} \times F_{TRN}^{C})]$ with
$D_{TRN}^{C} = D \setminus D_{TRN}$ and $F_{TRN}^{C} = F \setminus F_{TRN}$.
